---
title: "Less is More Expensive: Bulk Buying and Inequality"
author: "E. Mallick Hossain"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    number_sections: yes
bibliography: reference.bib
fig_caption: true
citation_package: natbib
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{amssymb, amsmath, bbm, booktabs, caption, etoolbox, hyperref, longtable, multirow, ragged2e, setspace, subcaption, xcolor}
  - \onehalfspace
  - \usepackage[multiple, symbol]{footmisc}
  # - \usepackage[nolists, noheads]{endfloat}
  # - \renewcommand{\efloatseparator}{\mbox{}}
  - \captionsetup[figure]{labelsep=colon,format=hang}
  - \captionsetup{font=footnotesize,skip=0pt}
---

\AfterEndEnvironment{tabular}{\tabularendstuff}
\newcommand{\tabularendstuff}{}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\begin{center}
\href{https://github.com/emallickhossain/WarehouseClubs/raw/master/code/6_paper/BulkBuyingInequality.pdf}{\textcolor{blue}{Updated Version Here}}\footnote[1]{I thank my advisers, Aviv Nevo, Holger Sieg, Katja Seim, and Sarah Moshary for their advice and constructive criticism. Thanks to Emek Basker, Chris Cronin, Frank DiTraglia, Eileen Divringi, Jim Ferry, Ishan Ghosh, Joao Granja, Ben Hyman, Jeff Lin, Paolo Martellini, Davin Reed, Paul Sangrey, Anna Tranfaglia, and Keith Wardrip for their comments. Thanks to Dmitri Koustas for providing data on warehouse clubs. The conclusions drawn from the Nielsen data are those of the researcher and do not reflect the views of Nielsen. Nielsen is not responsible for, had no role in, and was not involved in analyzing and preparing the results reported herein.}
\end{center}
 
\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

\begin{abstract}
\singlespace
I document that high-income households buy in bulk more often than low-income households, especially for storable, non-food items. By buying in bulk at the same rate as high-income households, low-income households could lower their grocery expenditures by five percent. These differences in bulk buying may bias standard measures of inequality and price indices. I focus on three factors that affect bulk buying: store access, cognitive constraints, and storage costs. Using new, hand-collected data on warehouse club entry and state-level price regulations, I find that warehouse clubs increase bulk buying only for middle- and high-income households while mandating the posting of unit prices, which reduce cognitive costs, increase bulk buying for all households. Using toilet paper purchases as a case study, I estimate that storage costs are important for all but the richest households and account for a large share of the bulk buying gap between high-income and low-income households. Counterfactual exercises show that mandating the display of unit prices would reduce the bulk buying gap by about one-third by encouraging households to take advantage of quantity discounts. Removing storage costs would close the remainder of the gap.
\end{abstract}

\newpage

# Introduction
Low-income households pay more than high-income households for similar products such as cars, appliances, and food [@fellowes2006]. Previous research shows that low-income households pay more for food due to differences in store access but home production helps households reduce food expenditures [@chung1999; @aguiar2005; @aguiar2007; @talukdar2008]. In this paper, I explore a novel hypothesis, that low-income households pay more than high-income households because they do not take advantage of quantity discounts. Using comprehensive data on household purchases and store-level assortments, I document that a bulk buying "gap" exists and that households can reduce their expenditures without sacrificing their consumption by using the quantity discounts currently available given their store and brand preferences. After establishing new reduced-form facts about the importance of *store access*, *cognitive constraints*, and *storage costs* to the bulk-buying decision, I embed these features into a demand model to discern which ones best explain why low-income households are less likely than high-income households to buy in bulk.

Households differ in their bulk buying, with households making over \$100,000 being about 30\% more likely to buy non-food items in bulk relative to observationally similar households making less than \$25,000 (i.e. same household composition, age, marital status, etc.).^["Bulk" sizes are packages in the top two quintiles of the size distribution for that product category. Non-food items are cleaning, household, and hygiene products like toilet paper, diapers, paper towels, etc.] If low-income households bought in bulk like high-income households, they could reduce their grocery expenditures by five percent and save an aggregate of \$2.7 billion.^[Throughout this paper, "high-income" refers to households making over \$100,000 and "low-income" refers to households making under \$25,000.]

Previous research highlights the value of quantity discounts [@nevo2009] and others document differences in bulk buying between income groups [@kunreuther1973; @beatty2010; @orhun2018]. However, these studies generally focus on a limited set of products and do not identify what factors generate these differences in bulk buying. I expand on this literature by examining bulk buying along the full range of *food* and *non-food* grocery products as well as exploring underlying factors that influence a household's decision to buy in bulk. I focus on three factors that contribute to this bulk-buying gap: *Store access*, *cognitive constraints*, and *storage costs*. 

First, *store access* could be a factor because stores do not stock all possible brands and sizes. For example, warehouse clubs, like Costco, stock large packages that are unavailable elsewhere.^[Retailer names throughout the paper are only for expository purposes and do not indicate whether particular retailers are present in the Nielsen data because retailer identities are masked.] Therefore, the stores that households shop at could affect the sizes they purchase. I show that low-income households are more likely to shop at dollar and discount stores while high-income households are more likely to shop at warehouse clubs. Using data on warehouse club entry, I find that middle- and high-income households increase their bulk buying by five to ten percent after a warehouse club opens nearby, while low-income households are unaffected.

However, *store access* is not the only contributing factor, because even *within* stores, high-income households are more likely to buy in bulk than low-income households. Using matched household and store-level data, I show that low-income households do not take full advantage of quantity discounts in stores where they currently shop. 

Second, *cognitive constraints* could affect whether or not households take advantage of quantity discounts. In particular, households may not realize that bigger packages offer a better per-unit value because it is costly to compute unit prices for all products. Put differently, unit prices may not be *salient* to the household's purchase decision. Research demonstrates that consumers respond to taxes when they are more prominently displayed, or more *salient* [@chetty2009; @finkelstein2009; @goldin2013]. Similarly, unit prices may not be salient unless they are presented in a clear and easily understood format. Using plausibly exogenous variation in state-level regulations mandating that unit prices be posted, I show that households are four to nine percent more likely to buy in bulk in states with unit price regulations.^[Throughout this paper, "unit price regulations" refer to rules regarding whether or not retailers must display unit prices.] By leveraging household moves between states with different regulatory regimes, I find that moving from a state without regulations to a state with unit price regulations increases bulk buying by three to four percent while moving from a state with regulation to a state without regulation does not have a significant effect on bulk purchasing. Overall, I provide strong evidence that bulk buying increases when unit prices are more salient and cognitive costs are reduced.

*Storage costs* are the third factor, which include the transportation and physical storage of the item, since both are directly related to an item's physical size. I compare the bulk buying gap across a wide range of products and document that "smaller" product categories, such as plastic wrap and detergent have smaller gaps than "larger" categories such as toilet paper and paper towels. These patterns suggest that storage costs are an important contributor to the bulk buying gap, but since the gap still remains for these smaller products, storage costs cannot be the sole contributing factor.

Other factors may also influence a household's decision to buy in bulk, such as liquidity constraints. These additional factors are not explored here, but @orhun2018 finds that liquidity constraints have a small effect on bulk buying.

My reduced-form evidence shows that changes in the retail environment (club entry and unit price regulations) affect bulk buying. I then estimate a demand model to predict how the bulk buying gap would change under different shopping environments such as implementing unit price regulations and reducing storage costs. 

To compute storage costs, I use a novel approach that does not rely on purchase histories, which is the standard approach [@hendel2006a; @hendel2006b]. Estimating storage costs supports the literature on second-degree price discrimination because firms use non-linear pricing to discriminate between different types of consumers [@pesendorfer2002; @nevo2002; @hendel2006a; @hendel2006b; @mcmanus2007; @stole2007; @cohen2008; @hendel2013; @hendel2014].^[There is a large macroeconomic literature that decomposes price dispersion into a variety of factors, but quantity discounts are often explicitly omitted. See @kaplan2015, @coibion2017, @kaplan2019, @hitsch2019 for discussions.] For storable goods, theory shows that quantity discounts enable firms to discriminate between households with high and low storage costs [@hendel2014]. I provide evidence that low-income households have high storage costs while high-income households have low storage costs. On balance, this means that firms extract higher rents from low-income households because they have less ability to store for the future. 

In my counterfactual exercises, I estimate that posting unit prices would reduce the bulk buying gap by about 30\% and reducing storage costs almost completely closes the gap. Since storage costs are so high, low-income households do not obtain much value from shopping at stores with wider assortments of sizes. If policymakers want to encourage households to buy in bulk, posting unit prices would reduce cognitive costs and increase bulk buying.

The rest of the paper is structured as follows. Section \ref{data} describes the data used for analysis. Section \ref{stylized} documents stylized facts of quantity discounting. Section \ref{factors} presents evidence of contributing factors to the bulk-buying gap. Section \ref{model} introduces the model. Section \ref{estimation} presents estimation results. Section \ref{counterfactual} shows the counterfactual exercises and Section \ref{conclusion} concludes.

# Data
```{r, echo=FALSE, eval=FALSE}
source("../1_getData/getCensusData.R")
source("../1_getData/getHomescan.R")
source("../1_getData/getZIPIncomePop.R")
```
In this section, I describe the datasets used for my analysis and give a brief overview of their respective features.^[Researcher's own analyses derived based in part on data from The Nielsen Company (US), LLC and marketing databases provided through the Nielsen Datasets at the Kilts Center for Marketing Data Center at The University of Chicago Booth School of Business.] Nielsen's Consumer Panel data provides information on household's shopping and purchasing decisions. Nielsen's Retail Scanner data provides information on weekly product assortments and prices. Combining the two datasets gives a comprehensive overview of each household's purchasing behavior and choice environment. I also use two new datasets on warehouse club locations and state-level regulations, which provide more information on households' shopping environment.

## Nielsen Consumer Panel Data
```{r, echo=FALSE, eval=FALSE}
source("../4_analyzeData/wrds/homescanSummaryStats.R")
```

I use the Nielsen Consumer Panel Dataset from 2004--2017. This dataset is a panel of about 178,000 unique households, of which I observe about 40,000 households each year from 2004--2006 and about 60,000 households each year from 2007--2017. Households scan all items that they purchase, input quantities, prices, date of purchase, and store. Nielsen retains about 80% of its panel from year to year with the mean and median tenure of a household being four and three years, respectively. 

I consider food, drink, and non-food grocery (e.g., paper towels, toilet paper, detergent, etc.) purchases made at grocery stores, discount stores, dollar stores, warehouse clubs, and drug stores. These outlets account for over 90\% of household expenditures in these categories. I exclude alcohol, tobacco, health, and general merchandise products from my analysis since these products may have different consumption patterns than food and cleaning products (e.g., cigarettes, painkillers) or are not suited for bulk purchases (e.g., printers, cookware, linens). I also exclude households with a student or military head of household as well as those with an annual income of less than \$5,000. Only about 3\% of households are excluded and I use the remaining 173,000 households for my analysis. See Appendix \ref{sec:sampleConstruction} for further details of sample construction.

Table \ref{tab:homescanSummaryStats} presents descriptive statistics for households in the sample. The following analysis uses Nielsen's projection weights to be nationally representative unless otherwise stated. Weights are computed to match moments based on household size, income, age, race, ethnicity, education, occupation, and presence of children.

\input{./finalTables/homescanSummaryStats.tex}

## Nielsen Scanner Data
The Nielsen Scanner Data contains weekly prices and volume sold of individual products at about 35,000 stores from about 90 retail chains between 2006--2016. I match the Retail Scanner Data with the Consumer Panel data based on store identification numbers and purchase dates. By matching the two datasets, I can recover both the household's choice set and the product it chose to purchase.

## Warehouse Club Data
I also use hand-collected data on all warehouse clubs in the United States between 2004--2015 gathered for @coibion2017. This data records information on the opening dates, locations, and identity of all warehouse clubs in the United States. It was gathered by combining information available on company websites, annual reports, and by contacting firms. 

## Unit Pricing Regulations
I compile a novel dataset on state-level regulations regarding the display of unit prices. The data is based on annual regulatory updates aggregated in Handbook 130 published by the National Institute of Standards and Technology [@nist130]. I cross-check this information with state regulatory codes and state officials to ensure accuracy. This data includes information on which states have regulations, when they were adopted, and how stringent these regulations are. More details are discussed in Section \ref{salience}.

# Stylized Facts
\label{stylized}

In this section, I document two stylized facts about the shopping landscape. First, I show that quantity discounts apply to 91\% of grocery categories. Second, I document that households making over \$100,000 are about 30\% more likely to buy non-food items in bulk than households making under \$25,000, compared to only three percent for food items. 

## Quantity Discount Prevalence
\label{prevalence}
<!-- This uses the bulkDiscount2016Scanner.R script -->
<!-- This uses the couponAndBulkSavings.R script -->
Simply put, quantity discounts are a specific form of non-linear pricing in which unit prices decrease as package size increases. To establish the prevalence and magnitude of quantity discounts, I use Nielsen's Retail Scanner data from 2016. I estimate quantity discounts using the following regression for each of the 693 product categories.^[38 categories could not be estimated typically because the data did not have sufficient variation. These were generally uncommon categories like mushroom sauce, canned grapes, and canned chow mein.]
\begin{equation}
\label{eq:bulkDiscount}
\ln(P)_{ibm} = \beta \ln(Size)_{ibm} + \lambda_{bm} + \epsilon_{ibm},
\end{equation}
where $P$ is the unit price (package price divided by package size) of product $i$ from brand $b$ purchased in market $m$ (defined as a store-week). $Size$ is the item's package size, which is the number of units included in a UPC (e.g., quart, square feet, count, pound, etc.). $\lambda$ is a brand-store-week fixed effect. Variation in unit prices across package sizes within the same brand-store-week identify $\beta$. If retailers offer quantity discounts, then $\beta$ will be negative.

Figure \ref{fig:bulkDiscountAllProds} plots the distribution of $\beta$ across product categories. I find that 91\% of all product categories have a statistically significant and negative $\beta$ and that non-food items generally have larger discounts than food items. The median $\beta$ is -0.53 for non-food products, which means that a 10 percent increase in package size is associated with a 5.3 percent decrease in unit price. This discount is larger than the median $\beta$ for food items (-0.43). The size and near-universality of quantity discounts suggest they offer substantial savings to households without sacrificing consumption.^[For a comparison of quantity discounts with coupons, see Appendix \ref{coupons}.]

\begin{figure}
  \centering
  \caption{Distribution of Bulk Discounts by Product Type}
  \includegraphics[height = 0.4\textwidth]{../5_figures/bulkDiscountAllProdsScanner.png}
  \caption*{Notes: Using Nielsen Retail Scanner data from 2016, this figure plots the distribution of coefficients from a regression of log unit price on log package size (Equation (\ref{eq:bulkDiscount})) for individual product categories. "Unit Price Change" denotes the percent reduction in a product's unit price for a one percent increase in the package size. Regression controls for store-brand-week fixed effects. Histogram plots 645 product categories.}
  \label{fig:bulkDiscountAllProds}
\end{figure}

## Bulk Purchasing
\label{bulkPurchasing}
<!-- This uses the discountingBehavior.R script -->
<!-- This uses the discountingBehaviorModules.R script -->
<!-- This uses the discountingBehaviorChannel.R script -->
Next, I document the take up of quantity discounts among different types of households. I classify a product as "bulk" if it is in the top two quintiles of the size distribution for that product category.^[This is the same definition as used in [@nevo2009].] Then, for each household, I compute the expenditure-weighted share of bulk purchases of food and non-food items. Thus, for each household, I have the share of purchases that were bulk purchases. I then regress this "bulk share" on household income and other household characteristics that could affect consumption patterns and may be correlated with income, and plot the income coefficients. The equation below is estimated on food and non-food purchases separately.
\begin{equation}
\label{eq:discountingBehavior}
BulkShare_{imt} = \sum_{q} \beta^{q} Income_{imt} + \gamma X_{imt} + \lambda_m + \lambda_t + \epsilon_{imt},
\end{equation}
where $BulkShare$ is household $i$'s share of bulk purchases in market $m$ in year $t$.^[I define a market as a Designated Market Area (DMA), which are non-overlapping groups of counties originally defined by Nielsen to measure television audience share. DMA's provide a finer geographic aggregation than MSA's while capturing markets that are bigger than single counties.] $Income$ consists of dummies for each income quantile $q$. $X$ consists of household demographics (age, household composition, marital status, and education).^[See Data Appendix for details of demographic variables and how they are collected.] Year and market fixed effects are captured by $\lambda$. Standard errors are clustered at the market level.

Figure \ref{fig:savingsBehavior} illustrates that bulk purchases compose a 11.5 percentage point larger share of non-food expenditures for households making over \$100,000 compared to those making \$5,000--\$8,000. As income increases, bulk purchases comprise an increasing share of expenditures. For food items, there is a more muted increase of one percentage point across income groups. 

The 11.5 percentage point gap is quite large. For the average household making between \$5,000-8,000, 34.8\% of their non-food grocery spending is on bulk packages. Hence, households making over \$100,000 are 33\% more likely to buy in bulk relative to the lowest-income group.^[When collapsing to income quartiles, the gap is slightly moderated, with households making over \$100,000 being 28\% more likely to buy in bulk than households making under \$25,000.]

\begin{figure}
\centering
\caption{Bulk Purchasing by Household Income and Product Type}
\includegraphics[height = 0.4\textwidth]{../5_figures/savingsBehavior.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income bin coefficients from Equation (\ref{eq:discountingBehavior}), which regresses the share of annual purchases that were bulk packages on income, household composition, age, marital status, and education, as well as market and year fixed effects. Nielsen projection weights are used to ensure national representativeness. Households making \$5--8k are the reference group. Standard errors are clustered at the DMA level. Coefficient values are reported in Appendix Tables \ref{tab:discountingBehaviorFood} and \ref{tab:discountingBehaviorNonFood}.}
\label{fig:savingsBehavior}
\end{figure}

These patterns are consistent with high-income households buying at the lowest unit price and consuming out of storage. Given the existence of quantity discounts, larger packages correspond to lower unit prices. The fact that low-income households are less likely to buy these storable items in bulk suggests that some obstacles may prevent them from buying and storing large packages. 

Across popular spending categories, these gaps are particularly large in storable, non-food categories like paper towels and toilet paper, where households making over \$100,000 are more than twice as likely to buy in bulk compared to households making under \$25,000. In popular food categories like milk and eggs, there is little relationship or even a negative relationship between income and bulk buying (See Figure \ref{fig:savingsBehaviorModules}).

\begin{figure}
\centering
\caption{Bulk Purchasing by Household Income and Product Type}
\includegraphics[height = 0.4\textwidth]{../5_figures/savingsBehaviorModules.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income bin coefficients from Equation (\ref{eq:discountingBehavior}), which regresses the share of annual purchases that were bulk packages on income, household composition, age, and marital status as well as market and year fixed effects. This regression is estimated for milk, eggs, diapers, toilet paper, and paper towels. Nielsen projection weights are used to ensure national representativeness. Households making \$5--8k are the reference group. Standard errors are clustered at the DMA level. Coefficient values are reported in Appendix Table \ref{tab:AppendixDiscountingBehaviorMods}}
\label{fig:savingsBehaviorModules}
\end{figure}

Since the bulk buying gap is largest for non-food products, the rest of this paper primarily focuses on non-food products. These products are ideal for analyzing bulk purchasing because they isolate the key features that make bulk buying and quantity discounts attractive for households. Primarily, households can store items for future consumption. Additionally, these products generally do not have substitutes and they cannot be produced at home (e.g., toilet paper, diapers, etc.). My findings carry over to food products, but one must be careful to account for perishability, which counteracts product storability. Additionally, many food products have close substitutes (e.g., soda, juice, water, etc.) and home production can substitute for many products [@aguiar2005; @aguiar2007].

## Savings from Bulk Buying
<!-- This uses the tpLowestPrice.R script -->
<!-- This uses the otherModsLowestPrice.R script -->
<!-- This uses the bulkDiscountSavingsEstimated.R script -->

In this subsection, I calculate the savings that households could achieve from buying in bulk using estimates from Sections \ref{prevalence} and \ref{bulkPurchasing} to predict savings if low-income households purchased the same size packages as high-income households. For each product category, I compute the average difference in package sizes by estimating the following regression:
\begin{equation}
\label{eq:discountingBehavior2}
\ln({AvgSize})_{imt} = \sum_q \beta^q Income_{imt} + \gamma X_{imt} + \lambda_m + \lambda_t + \epsilon_{imt},
\end{equation}
where $AvgSize$ is the size-weighted average package size purchased by household $i$ in market $m$ in year $t$.^[Average package size is weighted by quantity to account for the fact that an unweighted average would favor small packages.] $Income$ is an indicator for a household's income quartile. $X$ consists of household demographics (household composition, age, marital status, education). Market and year fixed effects are included through $\lambda$. 

In this regression, $\beta^q$ gives the average difference between the package size purchased by a household in income quartile $q$ and the lowest-income household (making less than \$25,000). To compute savings, I multiply this average difference in package size purchased by the quantity discount estimated in Section \ref{prevalence}. For example, high-income households buy 30\% larger package of toilet paper which has a quantity discount of 0.216. Therefore, low-income households could save $0.3 * 0.216 = 0.0648$ or about 6\% from buying big packages like high-income households do. Aggregating across all categories where high-income households buy larger packages gives an estimated savings of 5\%.^[This averages only across categories where high-income households buy larger packages. There are some categories, such as septic tank cleaners, clothespins, and soda which high-income households buy in smaller packages. Imposing that low-income households buy the same average size across *all* categories reduces projected savings to 2.3\%.] ^[See Appendix \ref{excessSpending} for an alternate method of computing savings using all choices available to households on their shopping trip.]

Saving 5\% on these common household purchases is substantial for low-income households. For the bottom quintile of the income distribution, these items account for 12\% of total household spending compared to 7\% for the top quintile of the distribution.^[Based on expenditure data on food at home and housekeeping supplies from Table 1 of the 2017 Consumer Expenditure Survey available at https://www.bls.gov/opub/reports/consumer-expenditures/2017/home.htm] If the about 11 million households making under \$25,000 were to obtain these savings, that would be an overall savings of \$2.7 billion. For context, this is equal to 4\% of the \$65 billion federal Supplemental Nutrition Assistance Program budget in 2018 [@snap2019]. Buying in bulk does not necessarily change how much households *consume*, it just changes how much they *buy* at one time.

# Factors Affecting Bulk Buying
\label{factors}

In this section, I provide evidence for the importance of *store access*, *cognitive constraints*, and *storage costs* to the bulk-buying gap. To do this, I use plausibly exogenous variation and natural experiments to estimate the causal impact of warehouse club entry and unit pricing regulation on bulk purchasing. 

## Store Access
The first potential contributor to the bulk buying gap is *store access*; low-income households may not live in areas where bulk sizes are available or may not shop at stores that offer bulk sizes. In this subsection, I provide evidence that the bulk buying gap persists within neighborhoods and within store types. Then, I show that warehouse club entry increases bulk buying by five to nine percent, but these increases hold only for middle- and high-income households.

### Inequality Within Markets and Retail Chains
<!-- This uses the sizeAvailabilityAllMods.R script -->
<!-- This uses the storeTypesByZip.R script -->
<!-- This uses the discountingBehaviorFE.R script -->

The first possibility is that store assortments differ between high- and low-income neighborhoods. This hypothesis is similar to the literature on food deserts, which analyzes whether supply differences contribute to nutritional inequality [@handbury2019]. To measure the average size offered by each store, I map each available package to its corresponding size quintile in its product category and then average across all products offered by the store. Figure \ref{fig:sizeAvailabilityAllMods} shows the relationship between the average size quintile offered in Retail Scanner stores and ZIP code median income. Within store types, there is little relationship between the average size offered and a ZIP code's median income. However, there are substantial differences in the sizes offered by different types of stores.

\begin{figure}
\centering
\caption{Average Package Quintile By Zip Code Median Income}
\includegraphics[height = 0.4\textwidth]{../5_figures/sizeAvailabilityAllModsChannelType.png}
\caption*{Notes: Using 2016 Nielsen Retail Scanner data, this figure plots the average size quintile offered within a store type within categorizes of ZIP code median income. The \$100,000 income bin includes all ZIP codes with incomes greater than \$100,000.}
\label{fig:sizeAvailabilityAllMods}
\end{figure}

Using ZIP Code Business Patterns data, Figure \ref{fig:storeTypesByZip} illustrates that there are substantial differences in the types of stores available in various neighborhoods. Richer ZIP codes have more large grocery stores per capita while poorer ZIP codes have more small grocery and dollar stores. Drug stores are more evenly distributed and discount and club stores tend to occupy middle-income ZIP codes.

\begin{figure}
\centering
\caption{Store Counts by ZIP Code Median Income}
\includegraphics[height = 0.4\textwidth]{../5_figures/storeTypesByZip.png}
\caption*{Notes: Using 2016 ZIP Code Business Patterns data, this figure plots the number of stores per 1,000 residents for the (population-weighted) average ZIP code in each income bin. The lowest income bin includes all ZIP codes with median incomes of \$30,000 or less and the highest income bin includes all ZIP codes with median incomes of greater than \$100,000. "Large" denotes a grocery store with 50 or more employees. The stores types correspond to the following 2012 NAICS codes: Discount and Club Stores -- 452910 (Warehouse Clubs and Supercenters); Grocery Stores -- 445110 (Supermarkets and Other Grocery (except Convenience) Stores); Dollar Stores -- 452990 (All Other General Merchandise Stores); Drug Stores -- 446110 (Pharmacies and Drug Stores).}
\label{fig:storeTypesByZip}
\end{figure}

If supply factors are the primary driver of the bulk buying gap, then the gap should disappear when comparing households in the same location since they have access to the same set of stores. I show that the bulk buying gap still persists within ZIP codes. This remaining gap corresponds to the amount that *cannot* be explained by differences in access, at least as approximated by geography.

The ideal experiment would randomly assign households to various shopping environments and compare the bulk buying behavior of each. However, the Nielsen data only reveals shopping choices after households have selected into neighborhoods. To the extent that households that live in the same neighborhood are unobservably similar, I interpret any differences between cross-sectional bulk buying and within-ZIP bulk buying as an upper bound on the share of the bulk buying gap that is attributable to store access.

Even within ZIP codes, there may be other factors affecting where households shop, such as whether or not a household has a vehicle, access to public transit, or a warehouse club membership. To account for possible differences, I examine how much of the bulk buying gap persists within chains. This exercise assumes that within a chain, households have access to the same assortment of goods [@dellavigna2019]. I also examine the bulk buying gap within store types (i.e. "channel") to account for the fact that bulk buying differences may primarily be between channels (discount versus dollar) instead of between retailers within a channel (Walmart versus Target).

I estimate within-ZIP and within-chain bulk buying gaps using a modified form of Equation \ref{eq:discountingBehavior}:
\begin{equation}
\label{eq:discountingBehaviorFE}
BulkShare_{imt} = \sum_q \beta^q Income_{imt} + \gamma X_{imt} + \lambda_{mt} + \epsilon_{imt},
\end{equation}
where $Income$ is an indicator for the income quantile of household $i$ in market $m$ in year $t$. $X$ consists of household demographics (household composition, age, marital status, and education). For the analysis of bulk buying within ZIP code, $BulkShare$ is the share of bulk purchases made by household $i$ in ZIP code $m$ in year $t$ and $\lambda_{mt}$ is a ZIP-year fixed effect. For the analysis of bulk buying within retail chains, $BulkShare$ is the share of bulk purchases made by household $i$ in retail chain $m$ in year $t$ and $\lambda_{mt}$ is a retail chain-year fixed effect and/or a channel fixed effect. 

Figure \ref{fig:discountingBehaviorFE} plots the income coefficients with and without fixed effects for each regression. Adding ZIP-year fixed effects reduces the gap between the highest and lowest income quartile by 5\% (from 10.1 percentage points to 9.6 percentage points). Results are unchanged if I use county-year fixed effects instead of ZIP-year fixed effects. Using channel-year fixed effects reduces the bulk buying gap by a more substantial 72\% (from 6.4 percentage points to 1.8 percentage points). Adding retail chain-year fixed effects on top of channel-year fixed effects does not significantly affect the bulk buying gap. This implies that a large share of the bulk buying gap is related to the *types* of stores households shop at, but not the specific chain they choose within a particular store type. 

\begin{figure}
\centering
\caption{Bulk Buying Within ZIP Codes or Stores}
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width = 0.9\textwidth]{../5_figures/discountingBehaviorFEZIP.png}
    \caption{ZIP-Year Fixed Effects}
  \end{subfigure}%
  \begin{subfigure}{0.5\textwidth}
    \centering
    \includegraphics[width = 0.9\textwidth]{../5_figures/discountingBehaviorFE.png}
    \caption{Retailer/Channel-Year Fixed Effects}
  \end{subfigure}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income bin coefficients from Equation (\ref{eq:discountingBehaviorFE}), which regresses the share of annual purchases that were bulk packages on income, household composition, age, marital status, and education as well as either ZIP code-year, retail chain-year, or store type-year fixed effects. Nielsen projection weights are used to ensure national representativeness. Households making \$5--8k are the reference group. Coefficient values are reported in Table \ref{tab:discountingBehaviorFEAll}.}
\label{fig:discountingBehaviorFE}
\end{figure}

Overall, within ZIP codes, the bulk buying gap between high- and low-income households persists. However, within store type (or retail chain), the bulk buying gap is substantially reduced. Two important conclusions can be drawn from these patterns. First, in an accounting sense, the type of store a household shops at accounts for almost three-quarters of the bulk buying gap. This is likely an overestimate given that there is endogeneity in store choice by households because many factors influence where households choose to shop. Second, the bulk buying gap *still persists* within channels and retail chains. These patterns suggest that where a household shops and what they choose within a store are much more important than where a household is located. The next section explores how store preferences are related to income and how warehouse clubs affect bulk buying.

### Store Preferences by Income
<!-- This uses the expendituresByChannel.R script -->

The previous section shows that while the bulk buying gap persists within ZIP codes, it is substantially moderated within store types and retail chains. In this section, I document differences in store preferences between income groups and then estimate the effect of warehouse club entry on bulk buying.

To demonstrate differences in store preference by household income, I examine the association of annual non-food purchase shares at various outlets while controlling for household characteristics:
\begin{equation}
\label{eq:storePreference}
ChannelShare_{imt} = \sum_q \beta^q Income_{imt} + \gamma X_{imt} + \lambda_{m} + \lambda_t + \epsilon_{imt},
\end{equation}
where $ChannelShare$ is the share of annual spending that household $i$ in market $m$ in year $t$ made in a particular channel (grocery store, discount store, dollar store, drug store, or warehouse club). $Income$ is an indicator for a household's income quantile. $X$ captures other household demographics (household composition, age, marital status, and education). Finally, market and year fixed effects capture differences in spending shares across markets and over time.

\begin{figure}
\centering
\caption{Annual Spending By Store Type, Relative to Low-Income Households}
\includegraphics[height = 0.4\textwidth]{../5_figures/expendituresByChannelColor.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income bin coefficients from Equation (\ref{eq:storePreference}), which regresses the share of annual purchases at each store type on income, household composition, age, marital status, and education as well as year and market fixed effects. Nielsen projection weights are used to ensure national representativeness. Households making \$5--8k are the reference group.}
\label{fig:expendituresByChannel}
\end{figure}

Figure \ref{fig:expendituresByChannel} reveals that while there are small differences in the share of annual expenditures at grocery, drug, and discount stores, there are dramatic differences in whether households shop at warehouse clubs or dollar stores: households making over \$100,000 spend about 13 percentage points more of their non-food expenditures at warehouse clubs than households making under \$25,000.

Since the biggest differences are in warehouse clubs and these stores almost solely stock bulk sizes, I focus on how warehouse clubs affect bulk buying. The first possibility is that high-income households shop at warehouse clubs because they are closer. Table \ref{tab:costcoDist} shows that low-income households are about 15 miles away from the nearest warehouse club compared to only 8 miles away for high-income households.

\input{./finalTables/costcoDist.tex}

The ideal experiment would randomly assign warehouse clubs to neighbornoods and then their effect on bulk buying could easily be calculated. However, store locations are not randomly assigned. Within a househod, it is exceedingly unlikely that a club opening could be co-incident with a shift in bulk buying, so any observed changes are likely causal. To determine whether households respond to warehouse club entry, I use hand-collected data on 419 warehouse club openings between 2004--2015.^[Data provided by the authors of @coibion2017 and covers BJ's, Costco, and Sam's Club.] Leveraging the panel structure of the Nielsen Consumer Panel, I estimate how a household's bulk purchasing changes after a warehouse club opens using the following equation:
\begin{equation}
\label{eq:costcoEntry}
BulkShare_{imt} = \beta Entry_{imt} + \gamma X_{imt} + \lambda_{im} + \epsilon_{imt},
\end{equation}
where $BulkShare$ is the share of bulk purchases made by household $i$ in market $m$ in quarter $t$. $Entry$ is an indicator for whether or not a club entered within 15 miles of household $i$ in quarter $t$.^[In cases where a household is located near multiple warehouse clubs, I use the earliest entry date since the first warehouse club would generate the largest supply shock.] ^[According to the 2017 National Household Travel Survey, the average household traveled about seven miles to buy goods, with low-income households traveling about one or two miles less than higher-income households [@nhts2017]. Allowing for the possibility that households might travel farther to shop at a warehouse club, I use a cutoff of 15 miles. Appendix Table \ref{tab:appendixCostcoEntryDifferentRadii} shows that this pattern is robust to other cutoffs.] I include a household-market fixed effect $\lambda$ to ensure that $\beta$ is identified by within-household changes in bulk buying before and after a club opens instead of households that may move to areas closer to warehouse clubs. $X$ controls for possible demographic changes within the household (household composition, age, marital status, and education). 

Table \ref{tab:costcoEntryDD} shows the regression results. Overall, households that experienced a warehouse club entry increased their bulk purchasing by about two percentage points. However, when I interact household income with club entry, the increase in bulk buying is due to changes for households making over \$25,000 and is increasing in income, with households making over \$100,000 increasing their bulk buying by four percentage points. Households in the lowest quartile do not have any significant change in their bulk buying. One likely reason that low-income households do not change their bulk buying is that even after a warehouse club enters, households do not purchase a membership (fees range from \$45-\$120 depending on the chain and membership level). Other possible reasons are that low-income households do not have access to transportation that can carry items home or even if they had a membership, they still would not purchase extremely large sizes available at warehouse clubs due to budget constraints.^[As an example, Philadelphia provides public transit access to a warehouse club. However, carrying club-sized items on a bus is infeasible for more than two or three items. A personal vehicle would be necessary.]

\input{./finalTables/costcoEntryDD.tex}

To illustrate the effect of club entry on bulk buying and check for pre-trends, I estimate Equation \ref{eq:costcoEntry}, but replace the entry indicator with dummies for each quarter pre- and post-entry:
\begin{equation}
\label{eq:costcoEventStudy}
BulkShare_{imt} = \sum_q \beta^q Qtr_{imt} + \gamma X_{imt} + \lambda_{im} + \epsilon_{imt},
\end{equation}
where $Qtr$ is a dummy for each quarter prior to entry and after entry, with the quarter immediately before entry ($q = -1$) as the reference group. Figure \ref{fig:eventStudyIncome} plots the quarterly coefficients and shows that there are no significant pre-trends and post-entry, there are significant increases in bulk buying for higher income households with households making over \$100,000 increasing their bulk buying by four to seven percentage points. These effects are persistent up to eight quarters after a warehouse club has opened.

\begin{figure}
\centering
\caption{Event Study of Warehouse Club Entry}
\includegraphics[height = 0.4\textwidth]{../5_figures/eventStudyIncome.png}
\caption*{Notes: This figure plots the quarterly coefficients from Equation (\ref{eq:costcoEventStudy})---the effects of warehouse club entry on bulk purchasing of households before and after club store entry---using 2004--2015 household-by-quarter Nielsen Consumer Panel data. The regression controls for household composition, age, marital status, and education as well as household-ZIP code fixed effects. All coefficients are relative to bulk purchasing in the quarter before entry ($q = -1$). Error bars denote 95\% confidence intervals.}
\label{fig:eventStudyIncome}
\end{figure}

This analysis estimates the intent to treat effect since not all households shop at the entrant warehouse club after it opens. As a result, this is a conservative lower bound on the actual treatment effect on households that shop at warehouse clubs.^[Even though low-income households do not change their bulk buying, other research suggests that they may be worse off because existing retailers are more likely to increase prices for storable products as a competitive response [@bauner2019].] The effect is quite substantial even given how conservative it is.

## Cognitive Constraints
\label{salience}
Cognitive constraints are another possible contributor to the bulk-buying gap. Consumers may not be aware of the quantity discount because they do not compute unit prices when making purchases. To test this hypothesis, I utilize a novel hand-collected dataset of state-level unit-price, which substantially reduce the cognitive costs of computing unit prices so households can compare products and pick the best value.

The emergence of unit price labeling dates back to the late 1960s and early 1970s. During this period, there was a large consumer protection movement that included a push for unit prices to be posted so consumers could compare different brands and sizes of products [@miyazaki2000]. While this provision was not incorporated into federal law, some retailers voluntarily posted unit prices. However, there was a lot of variation in how retailers calculated and displayed unit prices, making comparisons difficult. Much of this heterogeneity still persists today [@consumerreports2015].^[See Appendix Figure \ref{fig:unitPrice} for examples.]

To make comparisons easier, some states passed laws standardizing how unit prices were calculated so that consumers could more easily compare products. For example, instead of having to figure out how a per-ounce price compares to a per-100mL price for milk, all milk prices had to be calculated for the same base unit. Many of these states also specified the how the prices should be formatted for readability and clarity [@rose2000].

Using annual regulatory updates published by the National Institute of Standards and Technology (NIST), I compile a time series of state-level regulations on unit pricing [@nist130]. For states with regulations, I cross-checked NIST's designation with state regulatory codes and consulted with state officials to ensure accuracy. Figure \ref{fig:unitPriceLaws} shows that, as of 2017, 16 states have regulations on the display of unit prices and 34 have no regulations.^[Summary statistics of these groups are reported in Appendix Table \ref{tab:summaryStatsUnitLaws}.]

\begin{figure}
\centering
\caption{Unit Price Regulations by State (2017)}
\includegraphics[height = 0.4\textheight]{../5_figures/unitPriceLaws.png}
\caption*{Notes: Using data from NIST Handbook 130, this figure plots whether or not a state has regulations in place governing the display of unit prices as of August 1, 2017.}
\label{fig:unitPriceLaws}
\end{figure}

If these regulations affect household decisions, then bulk buying should differ between states with and without these regulations. To test this hypothesis, I estimate the differences in bulk purchasing between states with and without unit pricing regulations using the following regression:
\begin{equation}
\label{eq:unitPrice}
BulkShare_{it} = \beta_1 Reg_{it} + \gamma X_{it} + \lambda_t + \epsilon_{it},
\end{equation}

where $BulkShare$ is the annual share of expenditures that were bulk purchases for household $i$ in year $t$. $Reg$ is an indicator for whether or not unit-price regulations are in effect. $X$ controls for household composition, income, age, marital status, and education. I control for time fixed effects through $\lambda$. If there are different patterns across income groups within states that have unit pricing regulations, then the interaction term will pick them up. Standard errors are clustered by state because these regulations are at the state level.

Since 2004, no state has modified its regulations on unit prices, so the coefficient on unit pricing regulation is identified from cross-sectional variation between states that have regulations and those that do not. Table \ref{tab:unitPriceLaw} reports the results. Columns (1) and (2) reveal that bulk purchasing is three to four percentage points higher in states with unit price regulations compared to states without unit price regulation, even after controlling for household demographics and year fixed effects. Column (3) demonstrates that unit price regulation corresponds to an increase of at about two percentage points in bulk buying for the lowest income households, and there is some evidence that this effect is increasing in income. However, estimating at this level of detail is likely under-powered, so I only focus on differences between states for the remaining estimations.

\input{./finalTables/unitPriceLaw.tex}

I then analyze these unit pricing regulations at a higher level of detail. In particular, the regulations vary across two dimensions: Posting and Formatting. Table \ref{tab:unitPriceLawTable} shows the breakdown of states along these dimensions. First, states can opt to have unit price posting be voluntary (seven states) or mandatory (nine states). Second, states can specify how unit prices are formatted when they are displayed.^[All states with these regulations standardize how unit prices are to be calculated, which is what makes the voluntary states different from states without regulations.] Formatting regulations specify features including minimum font sizes, background colors, and label positioning. With the exception of California, only states that mandate unit price posting have formatting requirements. Excluding California, regulations are naturally ordered: no regulation, voluntary posting, mandatory posting (no formatting requirements), and mandatory posting (with formatting requirements). 

\input{./finalTables/unitPriceLawTable.tex}

Columns (4) and (5) continue the earlier analysis, but leverage the stringency of the regulations. Column (4) shows that mandatory posting is associated with significantly higher bulk buying, but states with voluntary requirements may have higher rates of bulk buying. However, as Table \ref{tab:unitPriceLawTable} shows, California is an outlier in this regulatory environment because is is the only state with the unique combination of voluntary posting and strict formatting requirements. Because of this, I exclude California and re-estimate the regression. Column (5) reveals that California is the primary driver of this effect and states with voluntary posting do not have significantly higher bulk purchasing. On the other hand, mandatory unit price posting is associated with a two to four percentage point increase in bulk buying. This pattern supports the intuition that standardized unit price presentation reduces cognitive costs, increase the *salience* of unit prices, and facilitate comparisons for consumers. 

This estimation provides strong evidence of a relationship between unit pricing regulations and bulk purchasing. However, there is a risk of selection bias since these regulations were primarily adopted in the Northeast and West Coast regions of the United States. Given the results, the selection would have to be such that states with households pre-disposed to buy in bulk were more likely to pass these regulations. This implication contradicts the narrative that these regulations were primarily focused on increasing consumer protection, so these states would likely have been ones focused on consumer protection and populations that may actually have been less likely to buy in bulk without this information provision. 

To avoid this possible selection issue and provide causal evidence, I examine the 2,219 households that move between states with different regulatory regimes. To estimate the effect of unit-price regulations on these movers, I use the following regression:
\begin{equation}
\label{eq:unitPriceMove}
BulkShare_{it} = \beta_1 Reg_{it} + \gamma X_{it} + \lambda_i + \lambda_t + \epsilon_{it},
\end{equation}
where the variables are the same as in Equation (\ref{eq:unitPrice}), but I control for household fixed effects and standard errors are clustered at the household level.^[Household fixed effects are actually household-income fixed effects to ensure that the interaction term is identified by differences in regulation and not by changes in a household's income.] With this specification, $\beta_1$ is identified by changes in bulk purchases for households that move from a state with unit-price regulations to a state without unit-price regulations (or vice versa).^[Projection weights are not used because the weights are not designed for this subsample of movers.] 

\input{./finalTables/unitPriceLawMovers.tex}

Table \ref{tab:unitPriceLawMovers} reports the results of estimating Equation (\ref{eq:unitPriceMove}). Columns (1) and (2) show that moving to a state with unit price regulation increases bulk buying by about one percentage point. This specification implicitly assumes that the effect of moving to a state with unit price regulations will be the same as moving to a state without regulations (i.e. the effect is symmetric). Column (3) treats the different directions of moving differently and shows that moving to a state without unit price regulations significantly decreases bulk buying, by 1.4 percentage points while moving to a state with regulations does not change bulk buying significantly. 

To illustrate this effect graphically, I plot an event study by estimating a modified version of Equation (\ref{eq:unitPriceMove}):
\begin{equation}
\label{eq:unitPriceEventStudy}
Y_{it} = \sum_T \beta_1^t E_{it} + \gamma X_{it} + \lambda_i + \lambda_t + \epsilon_{it},
\end{equation}
where $E$ is a dummy for each year before or after a household moves to a state with a different unit pricing regime. The reference group is $t = -1$ so all effects are relative to the year before the household moves. I estimate this equation separately for different types of moves and plot the annual coefficients in Figure \ref{fig:unitPriceEventStudy}. Moves from states without regulations to states with regulations may have different effects than moves from states with regulations to states without regulations. Figure \ref{fig:unitPriceEventStudy} shows that households decrease their bulk buying by about two percentage points when they move from a state with unit-price regulations to a state without unit-price regulations. On the other hand, households that move from states without unit-price regulations to states with unit pricing regulations do not change their bulk buying.

\begin{figure}
\centering
\caption{Event Study of Movers and Unit Price Regulations}
\includegraphics[height = 0.4\textwidth]{../5_figures/unitPriceEventStudy.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the $\beta_1^t$ coefficients and 95\% confidence intervals from Equation (\ref{eq:unitPriceEventStudy}), which regresses household bulk purchasing on dummies for years before and after a household moves to a state with a different unit pricing regime than the state it moves from. The regression controls for household composition, age, marital status, and education as well as household and year fixed effects. Standard errors are clustered at the household level. "To" reports estimates for households that move from a state without unit price regulations to a state with unit price regulations. "Away" reports estimates for households that move from a state with unit price regulations to a state without regulations.}
\label{fig:unitPriceEventStudy}
\end{figure}

Overall, this shows that unit-price regulations, which reduce cognitive costs and increase the *salience* of unit prices have an effect on bulk purchasing. The effect is similar to warehouse club entry, but more equitably spread across households. Since unit-price regulation increases bulk buying for all households with slightly larger increases for high-income households, this is strong evidence that improving education and messaging around understanding unit prices and how to utilize them will encourage more households to buy in bulk and lower their total shopping bill. 

Unit pricing regulations are relatively simple to implement for both policymakers and retailers. Retailers will bear some initial setup costs of redesigning their price labels, but ongoing costs will likely be similar to current menu costs that firms bear. Adopting unit pricing policies (like those recommended by the National Conference on Weights and Measures) would encourage bulk buying while imposing few costs. These findings support the broader assertion that increasing price transparency allows households to improve their decision-making.

## Storage Costs
Another contributor to the bulk buying gap is *storage costs*, which include both the cost of transporting the item home and the cost of storing it at home since both are related to the physical size of the item. If storage costs are the primary driver of the bulk buying gap, then smaller products should have smaller gaps than larger products. To test this, I estimate Equation \ref{eq:discountingBehavior2} (reproduced below) that relates average package sizes with household income:
\begin{equation*}
\ln(AvgSize)_{imt} = \sum_q \beta^q Income_{imt} + \gamma X_{imt} + \lambda_m + \lambda_t + \epsilon_{imt},
\end{equation*}
where $AvgSize$ is the average package size purchased by household $i$ in market $m$ in year $t$. $Income$ consists of dummies for each income quantile $q$. $X$ consists of household demographics (age, household composition, marital status, and education). Year and market fixed effects are captured by $\lambda$. 

Figure \ref{fig:discountingBehaviorChannelNonFood} plots the income coefficients from the regression for all non-food grocery categories. I have highlighted some popular product categories. The bulk buying gap is largest for the physically biggest products such as paper towels and toilet paper while the gap is smaller for less bulky items such as liquid detergent and plastic wrap. This pattern supports the hypothesis that storage costs contribute to the bulk buying gap, but the persistence of the gap even for smaller products suggests that other factors are at play.^[See Appendix Figure \ref{fig:discountingBehaviorChannelNonFoodMaxSize} for a similar analysis using the largest package size purchased by a household.]

\begin{figure}
\centering
\caption{Bulk Buying Gap For Non-Food Grocery Products}
\includegraphics[width = 0.9\textwidth]{../5_figures/discountingBehaviorChannelNonFood.png}
\caption{Average Package Size}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, these figures plot the $\beta$ coefficients from Equation (\ref{eq:discountingBehavior2}), which regresses average package size purchased on household income. The regression controls for household composition, age, marital status, and education as well as market and year fixed effects. Values are reported in Appendix Table \ref{tab:appendixDiscountingBehaviorChannelNonFood}.}
\label{fig:discountingBehaviorChannelNonFood}
\end{figure}

# Model

The previous analyses show that store access (especially warehouse clubs), cognitive constraints, and storage costs all affect bulk buying. To decompose the contribution of each, I embed these factors into a discrete choice model of the household's purchase decision. The ideal setting would include a homogeneous good available in a range of prices and sizes across different regulatory regimes. Differences in large and small purchases between high-income and low-income households would identify storage costs and differences in buying between regulatory regimes would identify cognitive constraints. This setting is approximated by one where products have limited dimensions of differentiation and demand is uncorrelated with income. 

A discrete choice model of toilet paper purchases closely approximates this ideal setting. Toilet paper is ideal for this analysis because it is a necessity item with easily observable dimensions of differentiation, namely price, quality, quantity, and package size. It is offered in a wide range of package sizes and stores stock numerous brands and sizes (grocery and mass merchandise stores usually stock 35--40 unique brand-sizes). The top five brands and private-label store brands account for 86\% of sales. I focus on the most common package sizes, which range from 4- to 24-roll packages. I define a product as a unique brand-size combination.^[Specifically, this is a unique brand-roll count-sheet count because packages can differ in their "concentration" due to "double," "mega," and "super mega" rolls.] Additionally, underlying toilet paper consumption is primarily a function of household composition and age, not income.^[A 100-fold cross-validated lasso regression of annual purchases on household characteristics rules out income as significantly predictive. See Appendix \ref{lasso} for details.] High-income households consume a similar amount as low-income households but make fewer purchases [@orhun2018]. Finally, toilet paper cannot be easily substituted for another product nor can it be obtained through home production.^[While a bidet is a possible alternative, this is more likely a lifestyle choice instead of a situation where households switch between toilet paper and bidets. Furthermore, in the United States, 98\% of households report that they use toilet paper (the remainder either said no or did not respond) [@statista2019].]

The biggest identification challenge is separately identifying storage costs from underlying demand (i.e. households may buy large quantities because they have high consumption or because they have low storage costs). To separate storage costs from demand, I use variation induced by differences in product concentration. Product concentration breaks the direct link between volume and consumption. In the detergent category, the same number of washes could be covered with a large diluted package or a smaller concentrated package. Both packages support the same underlying demand, but households with higher storage costs will prefer the smaller, more concentrated package over the larger one.

The same reasoning holds true for toilet paper. Households do not demand a particular number of rolls (the primary determinant of package size), but choose how long they want their supply to last (i.e. purchase enough to last for two weeks, a month, two months, etc.).^[According to a 2007 Charmin survey, the average person uses 57 sheets per day. I assume this consumption rate when computing how long a product will last [@jaffe2007].] Toilet paper comes in a variety of concentrations with "mega" rolls being more concentrated than "regular" rolls. Therefore, households that purchase 24 "regular" rolls have the same demand for toilet paper uses as households that purchase six "mega" rolls, but the former household has lower storage costs since they can store the bigger package.

To illustrate the varying concentrations of toilet paper, Figure \ref{fig:tpSizeQuantityDistributionAllProds} plots the distribution of quantity (measured in number of days the supply will last for a single person) against package sizes (measured in rolls) for toilet paper products in the Nielsen data. As expected, there is an increasing relationship between the number of rolls in a package and how long the package will last, but there is substantial variation within packages containing the same number of rolls. The dashed lines denote the 25th and 75th percentiles of the average days' supply purchased by households. A wide range of package sizes fall within this range for each brand.^[Scott toilet paper is an exception because it does not offer different roll types. All rolls have 1000 sheets.] For example, a household demanding a 60-day supply of Charmin could purchase a package containing anywhere from 8 to 24 rolls. This overlap generates the necessary variation to separate storage costs from underlying demand. 

\begin{figure}
\centering
\caption{Scatter-plot of Toilet Paper Package Size and Quantity}
\includegraphics[height = 0.4\textwidth]{../5_figures/tpSizeQuantityDistributionAllProds.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the package sizes and quantities of the top five toilet paper brands and private-label products. The x-axis represents the number of toilet paper rolls contained in a package while the y-axis represents the number of days a product will last a single person household assuming a consumption rate of 57 two-ply sheets per day. Noise is added horizontally to better illustrate the number of products available within package sizes since roll counts are discrete. Dashed lines indicate the 25th and 75th percentiles of the average days' supply purchased by a household.}
\label{fig:tpSizeQuantityDistributionAllProds}
\end{figure}

## Model Setup
I model a household's purchase decision using a static discrete choice framework. When making a purchase, households consider the unit price, quality, and size of each package and choose the package that maximizes their utility. These features are captured in the household's indirect utility function:
\begin{equation}
\label{eq:mlogit}
U_{ijt} = \underbrace{\beta_{1i} UnitPrice_{jt} + \beta_{2i} BigPack + \beta_{3i} SmallPack + \theta_{b(j)}}_{\beta' x_{ijt}} + \epsilon_{ijt},
\end{equation}
where $UnitPrice$ is the unit price of product $j$ at time $t$, $BigPack$ is a dummy for the package having more than 12 rolls, $SmallPack$ is a dummy for less than 12 rolls, and $\theta$ is a brand fixed effect.^[Unit prices are per 100 two-ply sheets.] ^[Households bunch at 12-roll packages, so this allows for different package preferences around this bunching point.] Brand fixed effects capture quality differences between products. 

To account for heterogeneity such as household demographics and state-level regulations that affect purchase decisions, I model $\beta$ as a function of these demographics (household composition, age, marital status, and education) and regulation. These enter by introducing interactions between household demographics and product characteristics. I assume $\epsilon$ is iid Type 1 extreme value. 

This simple model incorporates the key features necessary to decompose the bulk buying gap. Preferences for package size (a measure of storage costs) are captured by $\beta_2$ and $\beta_3$ while the salience of unit pricing is captured by the interaction of price and a dummy for unit-pricing regulations. 

The unit price coefficient is identified using price variation across shopping trips due to shopping at different stores or sales. The size coefficient is identified by variation in the product "concentration" as illustrated in Figure \ref{fig:tpSizeQuantityDistributionAllProds}. That is, given their preferred days' supply (y-value), some households choose large packages and some choose small packages (x-value). 

Given these assumptions and the structure of the error term, the probability that household $i$ chooses product $j$ on trip $t$ has a closed form:
\begin{equation}
P_{ijt} = \frac{e^{\beta' x_{ijt}}}{\sum_j e^{\beta' x_{ijt}}}
\end{equation}
As a result, the log-likelihood function can be written as:
\begin{equation}
LL(\beta) = \sum_t \sum_i \sum_j y_{ijt} \log(P_{ijt}),
\end{equation}
where $y$ indicates whether household $i$ chose product $j$ on shopping trip $t$. The preference parameters $\beta$ can then be estimated using MLE. 

# Estimation Results
\label{estimation}

I estimate this model separately for each income quartile and year. The coefficients for 2016 are reported in Table \ref{tab:mlogit2016}. I observe about 45,500 toilet paper purchases across about 14,800 households at grocery and mass merchandisers. 

\input{./finalTables/mlogit2016.tex}

The estimation results show that the unit price coefficient is negative, implying that all else equal, households prefer lower unit prices with lower income households being more price sensitive than high-income households. For all households, larger households tend to be more unit-price sensitive which matches intuition because, holding income constant, larger households have fewer resources are available per person. Notably, all households except for the richest are more price sensitive when unit price regulations are in effect. These results show that unit price regulations increase the salience of unit prices primarily for low- and middle-income households.

Figure \ref{fig:elasticity} plots the distribution of unit price elasticities (own-price) for each brand. The majority of elasticities fall between -1 and -3 with poorer households having larger elasticities (in magnitude).^[Table 4 of @cohen2008 reports elasticities ranging from -1.94 to -2.54 for paper towels. My estimates cover this range, but have a large mass between -1 and -2. Demand for toilet paper is likely less elastic than toilet paper since kitchen towels or paper napkins can substitute for paper towels. Toilet paper does not have any similar, readily available, substitutes.]

\begin{figure}
\centering
\caption{Distribution of Unit-Price Elasticity by Household Income}
\includegraphics[height = 0.4\textwidth]{../5_figures/elasticity2016.png}
\caption*{Notes: Using 2016 Nielsen Consumer Panel and Retail Scanner data, this figure plots the distribution of unit price elasticities resulting from the estimation of Equation \ref{eq:mlogit}.}
\label{fig:elasticity}
\end{figure}

Turning to storage costs, households making below \$100,000 are less likely to choose large package sizes, holding everything else constant. On the other hand, high-income households have no significant response to large package sizes. However, there are important differences between the storage costs of the lower quartiles. For households making less than \$25,000, they have a negative coefficient for large packages and a positive coefficient for small packages. These coefficients indicate a clear ordering where the poorest households prefer smaller packages to larger packages and are willing to pay accordingly. Converting these coefficients to a willingness-to-pay (WTP) measure (dividing the coefficient by the unit price coefficient), households would have to be compensated $0.961 / 6.640 = \$0.14$ per 100 sheets to purchase the large package and they would be willing to pay $0.711 / 6.640 = \$0.11$ per 100 sheets to obtain a smaller package. For context, the average unit price of toilet paper ranges from about \$0.15 (for Scott) to \$0.55 (for Cottonelle). Therefore, even for the cheapest brand, low-income households are willing to pay almost about 70\% more for a smaller package relative to a 12-roll package. Adding in unit price regulations makes them more price sensitive and lowers their necessary compensation slightly to $0.961 / (6.640 + 0.919) = \$0.13$ per 100 sheets to accept a large package. 

For households in the middle income quintiles, the pattern is slightly different. They prefer 12-roll packages (the omitted size) to large packages and are relatively indifferent to (or have a slight preferences against) small packages. Their WTP is \$0.16-\$0.24 per 100 sheets to avoid the large packages and \$0.00-\$0.03 per 100 sheets to avoid the small packages. Adding unit price regulations lowers their necessary compensation to \$0.15-\$0.19 per 100 sheets to accept the large package.

The richest households (those making over \$100,000) have a much lower (in magnitude) price sensitivity and do not significantly differ in their preferences for 12-roll packages and packages larger than 12 rolls. The relative indifference to large packages supports the hypothesis that they have much lower storage costs than low-income households.

These results support the earlier reduced-form findings that unit price regulations affect a household's bulk buying decision. Storage is costly for low-income households while high-income households have no discernible storage costs. These estimates show that low-income households are more price-sensitive, but their high storage costs counteract this preference. In the counterfactuals, I predict how the bulk buying gap changes in response to regulatory changes and reductions in storage costs.

## Model Fit
Table \ref{tab:modelFit} compares the overall model predictions to the actual data. My model captures the fact that high-income households buy more than low-income households, but it over-predicts the amount all households buy. This over-prediction is more severe for lower-income households and consequently, narrows the gap between high and low-income households. Since my model does not include individual product fixed-effects, it does not capture the fact that some products are disproportionately popular (or unpopular) relative to what would be expected based on their characteristics. For example, a particular Charmin 6-pack has a 9-10\% share for each income group, but based on its characteristics, the model only predicts a 5-6\% share. Including a product-specific effect would ensure a better fit, but at the cost of reducing the interpretability and intuition of the model. I opt to maintain the parsimony and interpretability of the model and simulate counterfactuals using this specification. 

\input{./finalTables/modelFit.tex}

# Counterfactuals
\label{counterfactual}

Using the parameter estimates from the previous section, I predict how households respond to lower storage costs and universal unit price regulation. For these counterfactual exercises, I compare all counterfactual results to a "base case" of predicted purchases given their current shopping environment. I consider two counterfactual scenarios:

1. **Unit-Price Regulation:** Unit-price regulations are adopted everywhere.

2. **Reduced Storage Costs:** All households have the same storage costs (i.e. size preferences) as high-income households.

For the unit-price regulation scenario, I set each household's price coefficient to the sum of their price coefficient and the regulation interaction term. For households making under \$25,000, their coefficient becomes $-6.640 + (-0.919) = -0.7559$. The other demographic interaction terms remain the same as before. For the reduced storage cost scenario, I set all size coefficients and their interactions equal to the coefficients for households making over \$100,000. 

Table \ref{tab:counterfactualMNLDays} reports the counterfactual predictions. Households making over \$100,000 do not change across counterfactuals because they do not have a significant response to unit price regulations and their size preferences are unchanged. However, households in the bottom two quintiles slightly increase the amount purchased when unit price regulations are put in effect, closing 31-32\% of the gap with the highest-income households. Removing storage costs closes another 50\% of the gap for households making between \$25,000-\$50,000 and the gap is completely reversed for households making under \$25,000 because they are buying more than the richest households. This reversal is primarily because they still are more price sensitive than high-income households. The model does a good job of capturing changes within income groups, but because it cannot account for disproportionately popular items, the between-income group changes and relationships are less robust. 

\input{./finalTables/counterfactualMNLDays.tex}

These counterfactuals support the main findings from Section \ref{factors} which showed that unit price regulations moderately increase bulk buying and that storage costs are a substantial factor preventing households from buying in bulk. Overall, reducing cognitive costs and increasing the salience of unit prices helps households make better value decisions, but storage costs are the primary factor preventing households from buying in bulk. 

# Conclusion

This paper documents the bulk-buying gap in which high-income households buy in bulk more often than low-income households, especially for storable, necessity items like toilet paper and paper towels. Because they buy small quantities and do not take advantage of quantity discounts, low-income households pay about 5\% more on their grocery bill. I provide evidence that *store access*, *cognitive constraints*, and *storage costs* contribute to this gap. 

I show that *where* a household shops accounts for a large portion of this disparity and that warehouse clubs increase bulk buying, but only for middle- and high-income households. Low-income households are unlikely to shop at warehouse clubs, even if they are nearby. Furthermore, using between-state variation in regulations on posting unit prices, I find that regulations that reduce cognitive costs and increase the *salience* of unit prices increase bulk buying.

Combining these features into a structural model of toilet paper purchases, I predict how households' bulk purchasing changes if unit-price regulations are adopted universally and if storage costs are removed. I find that unit-price regulations close the bulk buying gap by about 30\% while removing storage costs almost completely closes the gap (and for the poorest households, they buy even more than high-income households).

This paper is one of the first to focus on consumer's take-up of quantity discounts and explore the factors that contribute to this decision. It provides evidence that differences in bulk buying are due to *store access*, *cognitive constraints*, and *storage costs*. Furthermore, this paper demonstrates that inequality persists even within everyday shopping trips because high-income households are more likely to take advantage of quantity discounts than low-income households. These differences have substantial financial consequences for the poorest households and are likely to generate systematic underestimates of consumption inequality if quantity discounts offset quality differences between products. Additionally, if the prices of large and small packages evolve differently, then households may experience substantial changes in their buying power. Future work will determine the extent to which inequality and inflation measures are underestimated because of quantity discounts.

# References

<div id="refs"></div>

<!-- \processdelayedfloats -->

\newpage
\appendix

<!-- ## CPI Products and Weights -->
<!-- Table \ref{tab:cpi} shows the CPI weights of storable product categories that often exhibit quantity discounts. These weights come from the Bureau of Labor Statistics Handbook of Methods, Chapter 17: The Consumer Price Index (Updated 2-14-2018). To be conservative in the importance of accounting for quantity discounts, I capture products that are shelf-stable and storable to varying degrees. -->

<!-- The CPI tries to adjust for quantity discounts in that the "first multiple-unit price is reported for use in the CPI." However, to the extent that there are systematic differences in multi-unit purchases, especially for storable goods, price estimates based on bulk sizes are likely to overestimate the price index for households that consistently buy in bulk and underestimate the price index for households that buy small quantities. -->

<!-- **Full CPI Basket:** Includes food and beverages, housing, utilities, apparel, transportation, medical care, recreation, and education. -->

<!-- **"Store" Shopping Basket:** This basket is constructed to reflect the set of goods commonly found at grocery/supercenter/general merchandise stores. Includes food and alcohol at home (9.024), household furnishings (3.341), apparel (3.343), pet products (0.659), recreational goods (1.5), and other goods (1.634). I exclude categories such as rent, utilities, energy, transportation, health care, education, and services (including food service). -->

<!-- \input{./tables/cpi.tex} -->


# Quantity Discounts and Coupon Savings
\label{coupons}
<!-- This uses the discountSavings.R script -->
<!-- This uses the couponSavings.R script -->

This section compares savings from quantity discounts to savings from coupons. To be conservative, I compare the savings from redeemed coupons (likely higher than the average savings of all coupons offered) to savings offered by quantity discounts (likely lower than quantity discounts actually redeemed). For each product purchased in the Consumer Panel data, households can input the value saved if they used a coupon. For each product category, I compute the average discount across all products in that category.

I then estimate quantity discount savings based on moving from a product in the second quintile to the fourth quintile of the size distribution. This leaves out small product sizes that may have high unit prices due to convenience (e.g., a 20-oz soda bottle at the checkout counter) and especially large sizes that may not be widely available at all stores. This range covers sizes that households are likely to consider when making their purchase decision.

Figure \ref{fig:couponBulkSavings} plots the distribution of coupon savings and estimated bulk savings for food and non-food products. Coupon savings are narrowly clustered with a median savings of 31\% for non-food products and 33\% for food products. Bulk discounts have lower median savings for non-food and food products of 27\% and 23\%, respectively, but are more widely dispersed, even exceeding 50\% savings for some non-food products.^[Smaller shifts, such as from the second to third quintile or third to fourth quintile generate smaller savings, but still preserve the long right tail primarily for non-food products.] Coupon savings are similar across product categories while there is substantial variation in quantity discounts with non-food products offering higher savings.

\begin{figure}
  \centering
  \caption{Percent Savings from Coupons and Bulk Discounts}
  \includegraphics[height = 0.4\textwidth]{../5_figures/couponBulkSavings.png}
  \caption*{Notes: Using 2004--2017 Consumer Panel and 2016 Retail Scanner data, this figure plots the distribution of savings from coupons and quantity discounts. For each coupon redemption, the percent savings are the ratio of the coupon value to the product's price. These savings are then averaged across all purchases in that product category. Bulk discounts are computed using coefficient estimates obtained from Equation (\ref{eq:bulkDiscount}) relating log unit prices to log package sizes. Bulk savings are the estimated savings obtained from moving from the second to the fourth quintile of the size distribution for each product category.}
  \label{fig:couponBulkSavings}
\end{figure}

# Alternative Calculation of Missed Quantity Discounts
\label{excessSpending}
An alternative way of calculating savings from quantity discounts is to calculate first-best savings obtained from purchasing the lowest unit-priced item available, since even high-income households may not be buying at the lowest unit price. I compute this by taking the difference between the unit price paid by each household and the lowest unit price available in the store, given the household's brand preference. I get this information through linking the Nielsen Consumer Panel with the Nielsen Retail Scanner data.

I compute the first-best savings a household could obtain for toilet paper, diapers, milk, and eggs using the following approach.^[Due to computing and data storage constraints, I cannot do this for more than a few categories.] First, for each shopping trip, I compute the lowest unit price the household could have paid given its brand and store choice in that week. The difference in unit prices relative to the unit price chosen is a household's first-best savings for that purchase. Then, to get the average savings for a household, I compute the expenditure-weighted average savings across all purchases for each household. Based on this measure, Table \ref{tab:excessSpending} reports average excess spending by income group, computed for a family of four.

\input{./finalTables/lowestPrice.tex}

Overall, households could save over 30\% by buying in bulk and low-income households could save even more. I estimate the differences in savings between households from the following regression:
\begin{equation}
\label{eq:excessSpending}
Y_{imt} = \sum_q \beta^q Income_{imt} + \gamma X_{imt} + \lambda_{mt} + \epsilon_{imt},
\end{equation}
where $Y$ is the excess spending of household $i$ in market $m$ in year $t$. $Income$ is the household's income quantile and $X$ consists of demographic measures (age, marital status, household composition, education). $\lambda$ is a market-year fixed effect. Table \ref{tab:lowestPrice} shows that low-income households miss out on two percentage points more savings than high-income households and the excess spending is primarily in non-food categories like toilet paper (36\% savings) and diapers (33\% savings) as opposed to food categories like milk (31\% savings) and eggs (17\% savings). Given the perishability of food items, these savings may not be realized if the product perishes before it can be consumed. 

\input{./finalTables/lowestPriceRegFin.tex}

Overall, low-income households could benefit substantially from buying in bulk and obtaining lower unit prices. Furthermore, these savings are likely to be more important for low-income households since the marginal utility of an additional dollar of savings is likely to be higher than for high-income households. This analysis also provides evidence that all households could benefit from purchasing at the lowest unit price.


# Annual Consumption Analysis
\label{lasso}
I show that income is not predictive of a household's toilet paper consumption rate first using basic OLS regressions. I then formalize the result using a 100-fold cross-validated elastic net regression to select the most predictive variables. If income and toilet paper consumption are related, then an OLS regression will extract the correlation.

First, I compute a household's daily consumption by aggregating the total number of sheets purchased by a household in a given year, excluding the final purchase of the year since it may not be consumed within the year. I divide this total by the number of days between the first and last purchase of the year to get a household's average daily consumption rate. This method avoids complications where end of the year inventory may be carried over to the following year or a household may start the year with some inventory.

Given a household's average daily consumption rate, I estimate an OLS regression of consumption on household characteristics:
\begin{equation}
\label{eq:annualConsumption}
Y_i = \beta X_i + \epsilon_i,
\end{equation}
where $Y$ is household $i$'s average daily consumption and $X$ is a vector of household characteristics. Figure \ref{fig:annualPurchasesTP} plots the income coefficients of an OLS regression including only income covariates and the coefficients when household composition, age, marital status, and education are included. The graph illustrates that after controlling for covariates that plausibly cause increased consumption, income is not significantly correlated with consumption.

\begin{figure}
\centering
\caption{Average Daily Consumption by Household Income}
\includegraphics[height = 0.4\textwidth]{../5_figures/annualPurchasesTP.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income bin coefficients from Equation (\ref{eq:annualConsumption}), which regresses average daily household toilet paper consumption on household demographics (household composition, age, marital status, and education). Average daily consumption is computed by dividing total quantity purchased in a year by the number of days a household was active in the panel.}
\label{fig:annualPurchasesTP}
\end{figure}

The above specification omits many other possible covariates that could be correlated with average daily consumption. When there are many possible variables that can be included, there is a risk of over-fitting. Elastic net regularization is a machine learning method that penalizes over-fitting and selects only the most predictive variables.

The elastic net solves the following minimization problem:
\begin{equation}
\min_{\beta} \Vert y - X \beta \Vert^2 + \lambda \left(\alpha \Vert \beta \Vert_1 + (1 - \alpha) \Vert \beta \Vert^2_2 \right),
\end{equation}

where $\Vert \cdot \Vert_1$ is the L1 norm and $\Vert \cdot \Vert_2$ is the L2 norm. The OLS estimate is the $\beta$ that solves the minimization problem with only the first term. The second term and third term provide penalties to shrink and select for the most predictive variables.

I set the mixing parameter $\alpha$ to be 0.5. When covariates are correlated in groups, lasso regression ($\alpha = 1$) tends to only select one and discard all other members of the group while ridge regression ($\alpha = 0$) tends to shrink correlated coefficients towards each other [@zou2005]. Since some of the possible covariates form natural groups (e.g., all income bins or all markets), I chose $\alpha = 0.5$ since this tends to include or exclude groups together.

I estimate a 100-fold cross-validated elastic net regression to select the most predictive covariates. The resulting estimates selects many household characteristics including household composition, age, marital status, and race, but excludes almost all income and geographic coefficients.^[Elastic net results are available upon request.]

<!-- # Toilet Paper "Concentration" -->
<!-- To provide more detail on why toilet paper has different "concentrations," this section provides the geometric intution.  -->

<!-- Recall that "concentration" refers to the amount of product per unit volume. Therefore, in the case of toilet paper, this will be the ratio of the length of the paper to the volume of the roll. Therefore, to calculate concentration, two values must be computed.  -->

<!-- First, the volume is straightforward since a roll of toilet paper can be modeled as a cylinder. Therefore, the formula for volume is simply $\pi r^2 h$. Computing the length of unspooled toilet paper is more complicated. -->

<!-- A toilet paper roll can be modeled as an Archimedean spiral (i.e. arithmetic spiral), which has the following polar representation: -->
<!-- \begin{equation*} -->
<!-- r = a + b \theta, -->
<!-- \end{equation*} -->
<!-- where $a$ is the starting point of the spiral (distance from the origin) and $b$ is the distance between successive turns of the spiral.  -->

<!-- Given this representation, the length of the spiral takes the following form: -->
<!-- \begin{equation*} -->
<!-- L = \int_a^b \sqrt((a + b\theta)^2 + b^2) d\theta -->
<!-- \end{equation*} -->

<!-- While this formula has a closed form, it is easier to numerically integrate to get the length of the spiral.  -->

<!-- A cardboard toilet paper tube has an inner radius of about 0.8 inches (this is $a$) and a sheet of toilet paper is about 4.5 inches tall (this is $h$). Additionally, I assume toilet paper has a thickness of 0.01 inches (this is $b$). With these values, I can then compute the concentrations of various rolls of toilet paper. For illustration, I will compute this for rolls ranging from 50 turns to 150 turns (only Concentration column is rounded to two decimal places).  -->

<!-- | Turns | Length  | Volume       | Concentration | -->
<!-- |-------|---------|--------------|---------------| -->
<!-- | 50    | 744.816 | 7.605 \pi    | 31.17         | -->
<!-- | 75    | 1487.33 | 10.81125 \pi | 43.79         | -->
<!-- | 100   | 2476.59 | 14.58 \pi    | 54.07         | -->
<!-- | 125   | 3712.58 | 18.91125 \pi | 62.49         | -->
<!-- | 150   | 5195.32 | 23.805 \pi   | 69.47         | -->

<!-- The key takeaways are that concentration is changing as the roll gets bigger. To get a better intuition, compare the 100-turn spiral with the 150-turn spiral. The length has more than doubled (~110\% increase) while the volume has only increased by 60\%. In reality, the ratios are actually much larger because toilet paper is compressible, which means that manufacturers can wind the paper more tightly and are able to achieve large increases in length with negligible changes in volume. Figure \ref{fig:scott} shows that this compressability has even been advertised going back at least 100 years. -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \caption{Scott Toilet Paper Ad from 1915} -->
<!-- \includegraphics[height = 0.4\textwidth]{../5_figures/Scott_ad_1915.png} -->
<!-- \label{fig:scott} -->
<!-- \end{figure} -->

<!-- A brief empirical study undertaken by the author found that between two 12-roll packages of Quilted Northern and Angel Soft, a 60\% difference in the number of sheets corresponded to an 8\% difference in physical volume.  -->

<!-- # Bulk-Buying Gaps Across Product Categories -->
<!-- \label{discountingAll} -->
<!-- To supplement the analysis in Section \ref{bulkPurchasing}, this section provides a comprehensive analysis of bulk purchasing across all product categories. The main paper classified each product as a "bulk" size or not in order to aggregate purchasing across diverse product categories. In this section, I estimate a modified form of Equation \ref{eq:discountingBehavior} which regresses log average package size on income and other demographics. Since I estimate this regression for each of the hundreds of product categories, using log package size provides a more informative view of the difference in purchasing behavior between different household types. -->

<!-- For each household, I compute the total volume and number of packages purchased over the year for each product category. Dividing the total volume by the number of packages gives the average package size purchased by that household. For example, if a household purchased 20 rolls of toilet paper across two packages, then the average package size purchased is 10 rolls per package. Therefore, I have the average package size purchased for each household-year-product category. To establish that high income is associated with larger average packages, I estimate the following regression for each product category: -->
<!-- \begin{equation} -->
<!-- \label{eq:bulkBuyingGap} -->
<!-- log(Y)_{imt} = \beta_1 Income_{imt} + \beta_2 X_{imt} + \lambda_{m} + \lambda_t + \epsilon_{imt}, -->
<!-- \end{equation} -->
<!-- where $Y$ is the average package size purchased by household $i$ in market $m$ in year $t$. I define a market as a DMA.^[DMAs are non-overlapping groups of counties that define television markets, but they also provide reasonable sub-MSA geographic regions for analysis.] $Income$ is the household's income bin. $X$ controls for other household demographics including household composition, age, marital status, and education. Finally, to capture changes over time within geographic regions, I include a market and year fixed effects $\lambda$.  -->

<!-- Each product category gives a set of coefficients. These coefficients are plotted in Figure \ref{fig:discountingBehaviorChannelAll}. For clarity, I only show product categories in which at least 5000 households made a purchase in a year (about 5--10\% of Nielsen households). The figure shows that there are wide differences in average package sizes purchased across a wide variety of products. For paper towels, aluminum foil, and kitchen trash bags, households making over \$100,000 buy 35\% larger packages than those purchased by households making under \$25,000. Furthermore, the differences between income groups are statistically significant. Households making over \$100,000 purchase larger packages than those making \$50--100,000 who in turn purchase larger packages than those making \$25--50,000.  -->

<!-- In about 60\% of products, higher income households purchase larger packages.  -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \caption{Bulk Purchasing Across Product Categories and Income Groups} -->
<!-- \includegraphics[height = 0.4\textwidth]{../5_figures/discountingBehaviorChannelAll.png} -->
<!-- \caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the income coefficients from Equation (\ref{eq:bulkBuyingGap}) which regresses the log average package size purchased on household income, size, age, presence of children, marital status, and education as well as market and year fixed effects. Each bar represents a different product category and only statistically significant coefficients (at the 5\% level) are plotted. Values are relative to households making <\$25,000 (the reference category). Only categories for which more than 5000 households made a purchase in a given year are plotted.} -->
<!-- \label{fig:discountingBehaviorChannelAll} -->
<!-- \end{figure} -->

# Unit Pricing History and Regulations
Posting of unit prices started during the consumer protection movement of the late 1960s and early 1970s. Supporters argued that unit prices would allow consumers to make better decisions by facilitating comparisons between brands and sizes of the same product. During this era, several large grocery chains began voluntarily posting unit prices and on January 1, 1971, Massachusetts became the first state to mandate posting of unit prices [@isakson1973]. However, Figure \ref{fig:unitPrice} illustrates that there was much heterogeneity in how unit prices were presented and how prominently the unit price was displayed. Research found that unit prices help households identify package shrinkage and avoid quantity surcharges where bigger packages have higher unit prices than smaller ones [@yao2016]. Earlier research found that middle- and high-income households are more likely to use unit pricing information compared to low-income households to lower their expenditures [@isakson1973]. However, most of the research on consumer response to unit pricing has been done through consumer surveys (i.e. self-reports of awareness and/or use of unit pricing) or limited field experiments at stores in small markets on a small subset of products [@russo1977]. To the best of my knowledge, no analysis has been conducted on a nationally representative sample of consumer purchases like the Nielsen Consumer Panel to determine consumer response to unit pricing.

\begin{figure}
\centering
\caption{Unit Pricing Across Retailers}
\includegraphics[height = 0.4\textwidth]{../5_figures/unitPrice.png}
\caption*{Notes: This figure illustrates the heterogeneity in unit pricing presentation across a wide range of retailers. Image reproduces Figure 1 from Miyazaki et. al (2000).}
\label{fig:unitPrice}
\end{figure}

To promote "uniformity in weights and measures laws," the National Conference on Weights and Measures (supported by the National Institute of Standards and Technology) publishes recommended regulations that states can adopt. One of these regulations, the Uniform Unit Pricing Regulation (UPR), standardizes how unit prices are displayed across all package sizes of products. As of October 2018, eight states have adopted the recommended regulation and 10 states have adopted some form of unit price regulations that are not based on the UPR. The remaining 32 states have no regulations or the UPR is only referenced as a guideline. 

States may have varying levels of stringency regarding the display of unit prices. In particular, the UPR as specified only applies to retailers that voluntarily choose to display unit prices. However, if a retailer does display unit prices, the UPR becomes effective and requires the following (excerpts from the UPR are below): 

\S 2 [T]he unit price of a particular commodity in all package sizes ... shall be uniformly and consistently expressed in terms of [weight, volume, count, or area].

\S 4 (a) The unit price shall be to the nearest cent when a dollar or more

\S 4 (b) If the unit price is under a dollar, it shall be listed to the tent of a cent or to the whole cent, but not both.

\S 6 (a) If different brands or package sizes of the same consumer commodity are expressed in more than one unit of measure (e.g., soft drinks are offered for sale in 2 L bottles and 12 fl oz cans), the retailer establishment shall unit price the items consistently.

Within a commodity, unit prices must be specified in the same units (e.g., for drinks, unit prices can be per ounce or per liter, but cannot be expressed in different units for different packages).

An important fact is that the UPR gives retailers discretion over whether or not to display unit prices. However, as of 2015, nine states (Connecticut, Maryland, Massachusetts, New Hampshire, New Jersey, New York, Oregon, Rhode Island, and Vermont) have more stringent rules that mandate the display of unit prices by retailers and then specifying how unit prices are calculated and displayed [@nist1181]. In general, the requirements around how unit prices are calculated and displayed closely mirror those outlined in the UPR, so the main difference is whether retailers can voluntarily display unit prices or if they are mandated to display unit prices.

# Data Appendix

## Consumer Panel Data
\label{sec:sampleConstruction}
The Nielsen Consumer Panel consists of about 40,000--60,000 US households that provide information on their shopping purchases using in-home scanners or Nielsen's mobile app. Panelists are geographically dispersed and demographically balanced. Households are recruited based on key demographic characteristics, primarily household size, income, age, education, presence of children, race, ethnicity, and occupation. In order to generate national averages, Nielsen assigns each household a projection factor.

Households are recruited through direct mail and online invitations. To incentivize households to remain in the panel, Nielsen provides monthly prize drawings, sweepstakes, points, and regular communication and support to panelists. Nielsen tries to ensure that incentive methods are non-biasing and regularly tests for its correlation with retention rates.

To ensure data quality, Nielsen filters out any households that are poor reporters and do not meet minimum spending thresholds based on their household size. All households in the sample meet this threshold for the full year.

Demographic variables are recorded and updated annually. For my analysis, I collapse some of the demographic variables into more aggregate categories. Household composition measures the number of adult men, adult women, and children under the age of 18 residing in the home. Marital status is an indicator for whether the head of household is married or not (I do not distinguish between single, divorced, or widowed). Presence of children is an indicator for whether or not there is at least one child under the age of 18 present in the household. Education is an indicator for whether at least one head of household completed college. Finally, age is the age of the head of household. In the case of two heads, I average the two ages.

To construct my analysis sample, I remove any households where the head of household is a student or a member of the military because these households likely have different living arrangements that are not representative of a typical household's decision (i.e. on campus housing or barracks are different than traditional homes and apartments). I also remove any households making less that \$5,000 and those that could not be geocoded based on their ZIP code and the 2017 Census Gazetteer which provides a correspondence between ZIP codes and the latitude and longitude of the centroid of the ZIP code. Finally, some households were dropped because they could not be matched to tract-level car ownership data. Table \ref{tab:homeScanClean} reports how many households were removed based on this cleaning procedure.

\input{./finalTables/homeScanClean.tex}

In the purchase data, I exclude alcohol, tobacco, pet items, health and beauty items, general merchandise, "magnet," and "deferred" product categories from my analysis. Alcohol and tobacco are excluded because of their addictive qualities, which may induce peculiar purchase patterns. For example, a smoker may choose to only buy one pack of cigarettes with the intention of quitting even though a full carton may deliver a better value. Pet items are excluded to focus on products intended for human consumption. I exclude health and beauty items and general merchandise because these products such as trash cans, printers, eye shadow, and antacids are unlikely to be bought in bulk or have irregular consumption patterns. "Deferred" categories are categories that Nielsen has stopped tracking, so to maintain a consistent sample of products, these are excluded from my analysis. Finally, "magnet" purchases are items which do not have a UPC codes such as fresh fruits and vegetables, deli counter items, or bakery items. Because these items are only recorded for a subset of Nielsen households and are not standardized, I also exclude them from my analysis. This process leaves me with 721 unique product categories. Since this paper focuses on bulk purchases, I also exclude 28 categories that have five or fewer sizes across all products.^[These excluded categories are: jelled aspic salad, sour cream sauce mix, canned roast beef, canned roast beef hash, retort pouch bags, prepared sandwiches, canned rice, canned dumplings, canned bread, frozen vegetables in pastry, frozen grapefruit juice, frozen grape juice, frozen orange juice, frozen cream substitutes, canned ham patties, bathroom accessory, packaged soap, borateem, dry starch, grease relief, bathroom brushes, miscellaneous brushes, thermometers, dustpans, feather dusters, laundry baskets, sanitary belts, gift package with candy or gum.] Overall, the products analyzed are common household staples including almost all food categories, basic toiletry items, and non-food essentials like toilet paper, soaps/detergents, and diapers. See Table \ref{tab:homeScanProducts} for summary statistics of the top 20 product categories by annual spending.

\input{./finalTables/homeScanProductsFinal.tex}

In order to compare sizes across different product categories, I assign each product to its quintile in the size distribution for that product category. I assign quintiles based upon the sample quintiles of product sizes to ensure that each quintile has 20\% of available products in its support. An alternative strategy would assign quintiles based on cutting the range of product sizes into equal intervals. However, in some product categories, this risks generating quintiles with sparse support when there is an especially large package available. As an example, consider eggs. Most packages contain 6, 12, or 18 eggs, but there are some products that offer up to 15-dozen eggs (180 eggs). Generating quintiles by cutting the available range into equal intervals would generate quintiles of 1-36, 37-72, 73-108, 109-144, 145-180 which would assign almost all packages to the first quintile and the fifth quintile. Using the sample quintiles generates a more even distribution ensuring better support of each quintile. 
For products with a narrow range of sizes, I assign the product to the minimum quintile. For example, over 60\% of egg products are dozens, which covers three quintiles. I assign all products with 12 or fewer eggs to the first quintile. 

# Implications for Inequality
<!-- This uses the priceDecompChannelType.R script -->
Since quantity discounts are disproportionately taken up by high-income households, consumption inequality is persistently understated. In this section, I quantify the extent of this underestimation. 

I quantify consumption inequality in terms of unit prices since this is a more natural way of capturing quantity discounts. This approach abstracts away from consumption inequality attributable to differences in quantity consumed. For normal goods, adding this dimension will serve to further increase consumption inequality since high income households will consume larger quantities of the goods, compounding the inequality generated by lower unit prices. Unit prices are a conceptually simple way to approach consumption inequality. In the absence of quantity discounting, unit price differences would logically be attributed to differences in quality, such as brand or store amenities.

However, quantity discounts complicate this measurement. Consider the following example: a homogeneous product has a quantity discount. If some households buy large packages with low unit prices ("bulk" buyers) and some buy small packages with high unit prices ("budget" buyers), then the quantity discount would generate a difference in the underlying unit price. Even more troublingly, the inequality would be of the wrong sign. If one did not account for the quantity discount and only assumed unit price differences reflected differences in quality, then one would conclude that the "budget" buyers were better off than the "bulk" buyers, which is untrue since the "bulk" buyers obtain the same good more cheaply.

In order to quantify how much consumption inequality is underestimated, I decompose the unit price of each product category into store, brand, and size components using the following regression (estimated separately for each product category and year):
\begin{equation}
\label{eq:hedonic}
\log(unitPrice)_{ibr} = \beta \log(pkgSize)_{ibr} + \lambda_b + \lambda_r + \epsilon_{ibr},
\end{equation}
where $unitPrice$ is the unit price of a product purchased by household $i$ of brand $b$ at retailer $r$. $pkgSize$ is the quantity contained within the package. Brand and retailer components are captured by $\lambda$. All coefficients are relative to the reference category which is the generic brand product sold at the most popular discount retailer.

Using the estimates from Equation (\ref{eq:hedonic}), I decompose each purchase into a brand, retailer, and size component. I then aggregate these components to the household level by taking the expenditure-weighted average of the components across all purchases made by a particular household. I then regress each of these components on household income and other demographics (household composition, age, marital status, and education) to determine which components contribute to the difference in unit prices paid by different income groups. 
\begin{equation}
Y_{it} = \beta Income_{it} + \gamma X_{it} + \lambda_t + \epsilon_{it},
\end{equation}
where $Y$ is the average brand/retailer/size component for household $i$'s purchases made in year $t$. $X$ consists of household composition, age, marital status, and education. Year fixed effects are captured by $\lambda$. The regression is weighted using Nielsen's projection weights. Figure \ref{fig:tpPriceDecomp} plots the income group coefficients for each component on all toilet paper purchases. First, the plotted points show that the average unit price paid increases with household income. The decomposition shows that this increase is overwhelmingly due to brand composition with a small, but significantly positive contribution from retailer composition. Excluding a residual component, households making over \$100,000 would pay 14\% higher unit prices based on brand and store choices (13\% from brand choice and 1\% from store choice). Bulk discounts are estimated to reduce unit costs by six percentage points. As a result, consumption inequality in this category would be understated by 43\% (unit prices for high-income households would be 8\% compared to the "true" difference of 14\%). Due to a residual factor, the measured difference in unit prices is 9\%, so the "true" difference is understated by a still-substantial 36\%.

\begin{figure}
\centering
\caption{Decomposition of Unit Price Differences}
\includegraphics[height = 0.4\textwidth]{../5_figures/tpPriceDecomp.png}
\caption*{Notes: Using 2004--2017 Nielsen Consumer Panel data, this figure plots the contribution of brand, store, and package size choices to the differences in unit prices paid for toilet paper for different income groups relative to households making less than \$25,000. Dots denote difference in average unit price paid by each income group relative to households making under \$25,000. See Appendix Table \ref{tab:tpPriceDecomp} for values.}
\label{fig:tpPriceDecomp}
\end{figure}

\input{./finalTables/tpPriceDecomp.tex}



Bulk discounts have large implications for price measurement. Most macro-level research has focused on differences in aggregate spending between households or between various product categories [@krueger2006; @heathcote2010; @attanasio2014; @aguiar2015]. Under the assumption that homogeneous goods have the same prices, increased expenditures directly correspond to increased consumption. However, as @kaplan2015 show, prices for homogeneous goods are not the same and this non-uniformity has severe consequences for measurement. Most recently, @coibion2017 prove that consumption inequality is overstated due to bulk buying because households make fewer shopping trips and therefore measured price dispersion increases even if underlying consumption does not.^[This dispersion results because the Diary portion of the Consumer Expenditure Survey only records the past 2 weeks of purchases. Bulk buying allows some households to make purchases during those 2 weeks and others will not have to purchase during those 2 weeks.]

Using hedonic regression, I find that unit prices are underestimated by about 10\% due to quantity discounts, but this underestimation varies widely across product categories. See Appendix for analysis.


Previous research has documented that this "spikiness" has strong implications for measuring consumption inequality [@coibion2017]. Furthermore, quantity discounts could have implications for inflation measurement as well. Given the prevalence of quantity discounts, if a basket price is computed using the largest package available, measured prices may underestimate the price faced by consumers that cannot afford that package, or buy the same amount in smaller shopping trips. For example, the BLS uses unit-prices when computing the price of the CPI market basket. Therefore, the choice of package size can strongly influence this price. If the prices of different package sizes fluctuate differently, quantity discounts could systematically affect our measurement of inflation. Further research would be necessary to determine if inflation rates differ across product sizes is the case and what the magnitude could be.^[The BLS specifically deals with multi-unit discounts buy using the "first multiple-unit price" as the price measured. However, the method for dealing with different package sizes is less well defined. Per correspondence with the BLS: "In general, price-per-unit calculations are the standard practice in calculating items with quantity changes. For example if a loaf of bread is \$1 and you a second loaf for \$0.50, you have two loaves for \$1.50, and then depending on the weight of the two loaves, you divide the \$1.50 by number of ounces and get a price per ounce. Due to sample rotation, there is no guarantee that any specific item and specific size  toilet paper, olive oil, etc.  will be in the sample at a given time. If it is, and the economic assistant cant find the specified size and has to substitute to a different size, price change is calculated on a per unit basis, as illustrated in the above example." Given that the BLS uses unit-prices to compute market basket prices, the choice of package size could strongly influence the computed price of the market basket.]


# Appendix Figures
\begin{figure}[!htbp]
  \centering
  \includegraphics[width = 0.9\textwidth]{../5_figures/discountingBehaviorChannelNonFoodMaxSize.png}
  \caption{Max Size Purchased}
  \label{fig:discountingBehaviorChannelNonFoodMaxSize}
\end{figure}

# Appendix Tables
\input{./finalTables/AppendixDiscountingBehaviorFood.tex}
\input{./finalTables/AppendixDiscountingBehaviorNonFood.tex}
\input{./finalTables/AppendixDiscountingBehaviorFEAll.tex}
\input{./finalTables/AppendixDiscountingBehaviorMods.tex}
\input{./finalTables/appendixDiscountingBehaviorChannelNonFood.tex}
\input{./finalTables/summaryStatsUnitLaws.tex}
\input{./finalTables/appendixCostcoEntryDifferentRadii.tex}
